---
title: "Assignment2"
author: "483670 & 630516"
date: "08/09/2022"
output:
  pdf_document:
      includes:
        in_header: "preamble.tex"
---

```{r setup, include=FALSE}
rm(list=ls())
if (!require("pacman")) install.packages("pacman")
pacman::p_load(ggplot2, stargazer, lmtest, plm, 
               sandwich, fixest, modelsummary, vtable, foreign, RCT, reshape2)

#set tidy code
knitr::opts_chunk$set(tidy.opts = list(width.cutoff = 65), tidy = TRUE)
```

```{r data, include=FALSE}
data <- read.dta("searchperiod.dta")
```

# Q1

The mean probability of having benefits in week 10 of those without search period is $73.6%$. For those with the search period treatment this is $57.2%$. The probability of still having benefits in week 30 is $54%$ for those without and $41.4%$ for those with search treatment. In the most naive estimation it seems that the search treatment decreases the probability of being on benefits both 10 and 30 weeks after the initial claim.
```{r}
prob10nosearch <- mean(data$benefits_week10[data$searchperiod==0])
prob10nosearch
prob10search <- mean(data$benefits_week10[data$searchperiod==1])
prob10search
prob30nosearch <- mean(data$benefits_week30[data$searchperiod==0])
prob30nosearch
prob30search <- mean(data$benefits_week30[data$searchperiod==1])
prob30search
```

# Q2
The balance table is provided is provided in table \ref{tab:balance} The potential points of concern for the balance between control and treatment groups are the differences in age, slight unbalance in the locations and the difference in shares of unknown education. Particularly those with unknown education could be problematic. 
```{r 2}
balance <- balance_table(data, "searchperiod")
knitr::kable(balance, caption = "Balance Table")

```

# Q3
Table $\ref{tab:reg}$ below provides the regression output for the effect of the search period on the probability of being on benefits after 10 and 30 weeks both without and with controls. The effect of the treatment on the 10 week mark decreases slightly in magnitude from the naive estimate. It remains significant though at $14.3$ percentage points. The treatment effect increases in magnitude for being on benefits at the 30 week mark from a 9.9 percentage point to 12.6 percentage point decrease in the probability. In both cases including covariates maintains an economic and statistically significant treatment effect.
```{r 3, results='asis'}

m10b <- lm(benefits_week10 ~ searchperiod + sumincome_12monthsbefore+sumincome_24monthsbefore+age + female + children + partner + period2 +period3+period4 +location2+ location3+ location4+ location5+ educ_bachelormaster + educ_prepvocational + educ_unknown + educ_vocational, data = data)
m10a <- lm(benefits_week10 ~ searchperiod, data = data)
m30b <- lm(benefits_week30 ~ searchperiod, data = data)
m30a <- lm(benefits_week30 ~ searchperiod + sumincome_12monthsbefore+sumincome_24monthsbefore+age + female + children + partner + period2 +period3+period4 +location2+ location3+ location4+ location5+ educ_bachelormaster + educ_prepvocational + educ_unknown + educ_vocational, data = data)

stargazer(m10a, m10b, m30a, m30b, column.labels = c("", "Covariates","", "Covariates"),type = "latex",
          title = "LPM for Benefits", header = FALSE, label = "tab:reg", keep = c("searchperiod"))
```

# Q4
All the bounds are included in the overview table in tables \ref{tab:week10} and \ref{tab:week30} It is important to note the direction of the treatment effect. We are looking for a negative treatment effect, meaning that when we refer to the lower bound (mathematically) we are referring to the most optimistic bound on the effect size. We will make this clear in each case. The effect sizes are also measured in decimal points and should be interpreted as percentage point changes to the probability of being on benefits at the  10 and 30 week mark respectively. The no assumption bounds for the effect sizes  are $[-0.595, 0.405]$ at the 10 week mark and $[-0.561, 0.439]$ at the 30 week mark. This means that the most optimistic estimate for week 10 is a decrease in the probability of being on benefits of $59.5$ percentage points and the least optimistic is an increase by $40.5$ percentage points.
```{r}
probSearch = mean(data$searchperiod)
ymin = 0
ymax = 1
```

```{r 4, results='asis'}
# Benefits after 10 weeks - No assumption
Noass10_LB = prob10search*probSearch - prob10nosearch * (1-probSearch) + (ymin + ymax) * (1-probSearch) - ymax
Noass10_UB = prob10search*probSearch - prob10nosearch * (1-probSearch) + (ymin + ymax) * (1-probSearch) - ymin

# Benefits after 30 weeks - No assumption
Noass30_LB = prob30search*probSearch - prob30nosearch * (1-probSearch) + (ymin + ymax) * (1-probSearch) - ymax
Noass30_UB = prob30search*probSearch - prob30nosearch * (1-probSearch) + (ymin + ymax) * (1-probSearch) - ymin
```


# Q5
Here we impose the restrictions of the Roy model. We assume that only those who got the treatment benefit from it. Alternatively, this means that those who didn't get the treatment would not benefit from it. For the 10 week mark the bounds become $[-0.4, 0.261]$. For the 30 week mark the bounds are $[-0.294, 0.189]$. 
```{r 5, results='asis'}
# Benefits after 10 weeks - Roy Model
Roy10_LB = -(prob10nosearch - ymin)* (1-probSearch)
Roy10_UB =  (prob10search - ymin) * probSearch

# Benefits after 30 weeks - Roy Model
Roy30_LB = -(prob30nosearch - ymin)* (1-probSearch)
Roy30_UB =  (prob30search - ymin) * probSearch
```

# Q6
In MTR, treatment is assumed to be always improve the outcome. Therefore, we introduce the assumption that if an individual had a search period then this decreases the probability that they are going to be on benefits after 10 or 30 weeks after the treatment. This means that the probability that anyone is still on benefits after 10 or 30 weeks after treatment is always lower than if they had not received treatment. The MTR generally only affects the lower bound, as it pushes up the treatment effect (the worst estimate becomes better). However, in our case a better treatment effect is a more negative number. This means it works on our upper bound instead. This is why the upper bound becomes zero under MTR and the lower bound remains unchanged. The bounds under MRT for the 10 week mark the bounds become $[-0.595, 0.0]$. For the 30 week mark the bounds are $[-0.561, 0.0]$. 

MTS normally affects the upper bound as it restricts the best estimate of the treatment, assuming that individuals assigned to the treatment have more favorable outcomes than individuals who were not assigned to the treatment.  In our case though this is the lower bound. We introduce the assumption that individuals who got a search period have generally higher chances finding a job within 10 (30) weeks than individuals who were not assigned to treatment. The bounds under MTS for the 10 week mark the bounds become $[-0.164, 0.405]$. For the 30 week mark the bounds are $[-0.126, 0.439]$.

Combining both assumptions yields the MTS/MTR bounds of $[-0.164, 0.00]$ and $[-0.126, 0.0]$ for the 10 and 30 year mark respectively.
```{r 6, results='asis'}
# Benefits after 10 weeks - MTS/MTR
MTS10_LB = prob10search - prob10nosearch
MTS10_UB = prob10search*probSearch - prob10nosearch * (1-probSearch) + (ymin + ymax) * (1-probSearch) - ymin

MTR10_LB = prob10search*probSearch - prob10nosearch * (1-probSearch) + (ymin + ymax) * (1-probSearch) - ymax
MTR10_UB = 0

MTSMTR10_LB = MTS10_LB
MTSMTR10_UB = MTR10_UB

# Benefits after 30 weeks - MTS/MTR
MTS30_LB = prob30search - prob30nosearch
MTS30_UB = prob30search*probSearch - prob30nosearch * (1-probSearch) + (ymin + ymax) * (1-probSearch) - ymin
  
MTR30_LB = prob30search*probSearch - prob30nosearch * (1-probSearch) + (ymin + ymax) * (1-probSearch) - ymax
MTR30_UB = 0

MTSMTR30_LB = MTS30_LB
MTSMTR30_UB = MTR30_UB
```

# Q7

```{r}

prob10search_bachelormaster = mean(data$benefits_week10[data$searchperiod==1 & data$educ_bachelormaster==1])
prob10nosearch_bachelormaster = mean(data$benefits_week10[data$searchperiod==0 & data$educ_bachelormaster==1])

prob10search_vocational = mean(data$benefits_week10[data$searchperiod==1 & data$educ_vocational==1])
prob10nosearch_vocational = mean(data$benefits_week10[data$searchperiod==0 & data$educ_vocational==1])

prob10search_prepvocational = mean(data$benefits_week10[data$searchperiod==1 & data$educ_prepvocational==1])
prob10nosearch_prepvocational = mean(data$benefits_week10[data$searchperiod==0 & data$educ_prepvocational==1])

# Goes against the MIV assumption!

prob10search_primaryorless = mean(data$benefits_week10[data$searchperiod==1 & data$educ_primaryorless==1])
prob10nosearch_primaryorless = mean(data$benefits_week10[data$searchperiod==0 & data$educ_primaryorless==1])

# Does not tell us anything, so ignore
# prob10search_unknown = mean(data$benefits_week10[data$searchperiod==1 & data$educ_unknown==1])
# prob10nosearch_unknown = mean(data$benefits_week10[data$searchperiod==0 & data$educ_unknown==1])
```

Checking the data for the assumptions of the MIV reveals that Vocational, Prepvocational and Primaryorless go against the MIV assumption. We thus reduce the MIV to whether a person has a bachelor or not, and check again whether the MIV assumption holds:
```{r}
prob10search_NObachelormaster = mean(data$benefits_week10[data$searchperiod==1 & data$educ_bachelormaster==0])
prob10nosearch_NObachelormaster = mean(data$benefits_week10[data$searchperiod==0 & data$educ_bachelormaster==0])
```

Now, the MIV assumption holds as for both cases $d = \{0,1\}$, the probability of persons with a bachelor or master degree being still unemployed after 10 weeks is lower than the probability of persons without a bachelor or master degree. Logically this could also be explained by the biggest difference in educational qualification existing for those with a bachelor and master degree. In contrast, the others may give people similarly marketable skills.
```{r}
prob_BachelorMaster = mean(data$educ_bachelormaster)
```

```{r}
# MIV for Benefits after 10 weeks

prob10nosearch_LB = (1-prob_BachelorMaster) * mean(data$benefits_week10[data$searchperiod==0 & data$educ_bachelormaster==0]) + prob_BachelorMaster * max(mean(data$benefits_week10[data$searchperiod==0 & data$educ_bachelormaster==0]), mean(data$benefits_week10[data$searchperiod==0 & data$educ_bachelormaster==1])) 
prob10nosearch_UB = (1-prob_BachelorMaster) * min(mean(data$benefits_week10[data$searchperiod==0 & data$educ_bachelormaster==0]), mean(data$benefits_week10[data$searchperiod==0 & data$educ_bachelormaster==1])) + prob_BachelorMaster * mean(data$benefits_week10[data$searchperiod==0 & data$educ_bachelormaster==1])

prob10search_LB = (1-prob_BachelorMaster) * mean(data$benefits_week10[data$searchperiod==1 & data$educ_bachelormaster==0]) + prob_BachelorMaster * max(mean(data$benefits_week10[data$searchperiod==1 & data$educ_bachelormaster==0]), mean(data$benefits_week10[data$searchperiod==1 & data$educ_bachelormaster==1])) 
prob10search_UB = (1-prob_BachelorMaster) * min(mean(data$benefits_week10[data$searchperiod==1 & data$educ_bachelormaster==0]), mean(data$benefits_week10[data$searchperiod==1 & data$educ_bachelormaster==1])) + prob_BachelorMaster * mean(data$benefits_week10[data$searchperiod==1 & data$educ_bachelormaster==1])

MIV10_LB = prob10search_LB - prob10nosearch_LB

MIV10_UB = prob10search_UB - prob10nosearch_UB
```

```{r}
# MIV for Benefits after 30 weeks

prob30nosearch_LB = (1-prob_BachelorMaster) * mean(data$benefits_week30[data$searchperiod==0 & data$educ_bachelormaster==0]) + prob_BachelorMaster * max(mean(data$benefits_week30[data$searchperiod==0 & data$educ_bachelormaster==0]), mean(data$benefits_week30[data$searchperiod==0 & data$educ_bachelormaster==1])) 
prob30nosearch_UB = (1-prob_BachelorMaster) * min(mean(data$benefits_week30[data$searchperiod==0 & data$educ_bachelormaster==0]), mean(data$benefits_week30[data$searchperiod==0 & data$educ_bachelormaster==1])) + prob_BachelorMaster * mean(data$benefits_week30[data$searchperiod==0 & data$educ_bachelormaster==1])

prob30search_LB = (1-prob_BachelorMaster) * mean(data$benefits_week30[data$searchperiod==1 & data$educ_bachelormaster==0]) + prob_BachelorMaster * max(mean(data$benefits_week30[data$searchperiod==1 & data$educ_bachelormaster==0]), mean(data$benefits_week30[data$searchperiod==1 & data$educ_bachelormaster==1])) 
prob30search_UB = (1-prob_BachelorMaster) * min(mean(data$benefits_week30[data$searchperiod==1 & data$educ_bachelormaster==0]), mean(data$benefits_week30[data$searchperiod==1 & data$educ_bachelormaster==1])) + prob_BachelorMaster * mean(data$benefits_week30[data$searchperiod==1 & data$educ_bachelormaster==1])

MIV30_LB = prob30search_LB - prob30nosearch_LB

MIV30_UB = prob30search_UB - prob30nosearch_UB
```

```{r}
# Dataframe with all bounds for benefits after 10 weeks
Model <- c('No Assumptions', 'Roy Model', 'MTS', 'MTR', 'MTS+MTR', 'MIV')
Lowerbound <- c(Noass10_LB, Roy10_LB, MTS10_LB, MTR10_LB, MTSMTR10_LB, MIV10_LB)
Upperbound <- c(Noass10_UB, Roy10_UB, MTS10_UB, MTR10_UB, MTSMTR10_UB, MIV10_UB)

bounds10 <- data.frame(Model, Lowerbound, Upperbound)
knitr::kable(bounds10, caption = "10 Week Bounds")

# Dataframe with all bounds for benefits after 30 weeks
Model <- c('No Assumptions', 'Roy Model', 'MTS', 'MTR', 'MTS+MTR', 'MIV')
Lowerbound <- c(Noass30_LB, Roy30_LB, MTS30_LB, MTR30_LB, MTSMTR30_LB, MIV30_LB)
Upperbound <- c(Noass30_UB, Roy30_UB, MTS30_UB, MTR30_UB, MTSMTR30_UB, MIV30_UB)

bounds30 <- data.frame(Model, Lowerbound, Upperbound)
knitr::kable(bounds30, caption = "30 Week Bounds")
```
```{r}
bounds10$Model  = factor(bounds10$Model, levels=c("No Assumptions", "Roy Model", "MTS", "MTR", "MTS+MTR", "MIV"))
d <- melt(bounds10, id.vars="Model", value.name = "TreatmentEffect")

# Everything on the same plot
ggplot(d, aes(Model, TreatmentEffect, col=variable)) + 
  geom_line( color="grey") +
  geom_point(shape="-", color="#69b3a2", size=15) +
  stat_smooth() +
  geom_hline(yintercept=0)
```

```{r}
bounds30$Model  = factor(bounds30$Model, levels=c("No Assumptions", "Roy Model", "MTS", "MTR", "MTS+MTR", "MIV"))
e <- melt(bounds30, id.vars="Model", value.name = "TreatmentEffect")

# Everything on the same plot
ggplot(e, aes(Model, TreatmentEffect, col=variable)) + 
  geom_line( color="grey") +
  geom_point(shape="-", color="#69b3a2", size=15) +
  stat_smooth() +
  geom_hline(yintercept=0)
```

