---
title: "Assignment2"
author: "483670 & 630516"
date: "08/09/2022"
output:
  pdf_document:
      includes:
        in_header: "preamble.tex"
---

```{r setup, include=FALSE}
rm(list=ls())
if (!require("pacman")) install.packages("pacman")
pacman::p_load(ggplot2, stargazer, lmtest, plm, 
               sandwich, fixest, modelsummary, vtable, foreign, RCT)
```

```{r data, include=FALSE}
data <- read.dta("searchperiod.dta")
```

# Q1

The mean probability of having benefits in week 10 of those without search period is $73.6%$. For those with the search period treatment this is $57.2%$. For 
```{r}
prob10nosearch <- mean(data$benefits_week10[data$searchperiod==0])
prob10search <- mean(data$benefits_week10[data$searchperiod==1])
prob30nosearch <- mean(data$benefits_week30[data$searchperiod==0])
prob30search <- mean(data$benefits_week30[data$searchperiod==1])
```

# Q2

```{r 2}
balance <- balance_table(data, "searchperiod")
balance
```

#Q3

```{r 3, results='asis'}

m10b <- lm(benefits_week10 ~ searchperiod + sumincome_12monthsbefore+sumincome_24monthsbefore+age + female + children + partner + period2 +period3+period4 +location2+ location3+ location4+ location5+ educ_bachelormaster + educ_prepvocational + educ_unknown + educ_vocational, data = data)
m10a <- lm(benefits_week10 ~ searchperiod, data = data)
m30b <- lm(benefits_week30 ~ searchperiod, data = data)
m30a <- lm(benefits_week30 ~ searchperiod + sumincome_12monthsbefore+sumincome_24monthsbefore+age + female + children + partner + period2 +period3+period4 +location2+ location3+ location4+ location5+ educ_bachelormaster + educ_prepvocational + educ_unknown + educ_vocational, data = data)

stargazer(m10a, m10b, m30a, m30b, column.labels = c("", "Covariates","", "Covariates"),type = "latex",
          title = "LPM for Benefits", header = FALSE)
```

#Q4

```{r}
probSearch = mean(data$searchperiod)
ymin = 0
ymax = 1
# we can also jsut keep it in decimals
```

```{r 4, results='asis'}
# Benefits after 10 weeks - No assumption
Noass10_LB = prob10search*probSearch - prob10nosearch * (1-probSearch) + (ymin + ymax) * (1-probSearch) - ymax
Noass10_UB = prob10search*probSearch - prob10nosearch * (1-probSearch) + (ymin + ymax) * (1-probSearch) - ymin

Noass_LB=57.2*0.4563-73.6*0.5437+100*0.53437-100
Noass_UB=57.2*0.4563-73.6*0.5437+100*0.53437

# Benefits after 30 weeks - No assumption
Noass30_LB = prob30search*probSearch - prob30nosearch * (1-probSearch) + (ymin + ymax) * (1-probSearch) - ymax
Noass30_UB = prob30search*probSearch - prob30nosearch * (1-probSearch) + (ymin + ymax) * (1-probSearch) - ymin
```


#Q5

```{r 5, results='asis'}
# Benefits after 10 weeks - Roy Model
Roy10_LB = -(prob10nosearch - ymin)* (1-probSearch)
Roy10_UB =  (prob10search - ymin) * probSearch
Roy_LB=-73.6*0.5437
Roy_UB=57.2*0.4563

# Benefits after 30 weeks - Roy Model
Roy30_LB = -(prob30nosearch - ymin)* (1-probSearch)
Roy30_UB =  (prob30search - ymin) * probSearch
```

#Q6

```{r 6, results='asis'}
# Benefits after 10 weeks - MTS/MTR
MTS10_LB = prob10search - prob10nosearch
MTS10_UB = prob10search*probSearch - prob10nosearch * (1-probSearch) + (ymin + ymax) * (1-probSearch) - ymin
  
MTS_LB=57.2*0.4563-73.6*0.5437+100*0.53437-100
MTS_UB=0

MTR10_LB = prob10search*probSearch - prob10nosearch * (1-probSearch) + (ymin + ymax) * (1-probSearch) - ymax
MTR10_UB = 0

MTR_LB=-16.4
MTR_UB=57.2*0.4563-73.6*0.5437+100*0.53437

MTSMTR10_LB = MTS10_LB
MTSMTR10_UB = MTR10_UB

MTSMTR_LB=-16.4
MTSMTR_UB=0

# Benefits after 30 weeks - MTS/MTR
MTS30_LB = prob30search - prob30nosearch
MTS30_UB = prob30search*probSearch - prob30nosearch * (1-probSearch) + (ymin + ymax) * (1-probSearch) - ymin
  
MTR30_LB = prob30search*probSearch - prob30nosearch * (1-probSearch) + (ymin + ymax) * (1-probSearch) - ymax
MTR30_UB = 0

MTSMTR30_LB = MTS30_LB
MTSMTR30_UB = MTR30_UB
```

#Q7

```{r}

prob10search_bachelormaster = mean(data$benefits_week10[data$searchperiod==1 & data$educ_bachelormaster==1])
prob10nosearch_bachelormaster = mean(data$benefits_week10[data$searchperiod==0 & data$educ_bachelormaster==1])

prob10search_vocational = mean(data$benefits_week10[data$searchperiod==1 & data$educ_vocational==1])
prob10nosearch_vocational = mean(data$benefits_week10[data$searchperiod==0 & data$educ_vocational==1])

prob10search_prepvocational = mean(data$benefits_week10[data$searchperiod==1 & data$educ_prepvocational==1])
prob10nosearch_prepvocational = mean(data$benefits_week10[data$searchperiod==0 & data$educ_prepvocational==1])

# Goes against the MIV assumption!

prob10search_primaryorless = mean(data$benefits_week10[data$searchperiod==1 & data$educ_primaryorless==1])
prob10nosearch_primaryorless = mean(data$benefits_week10[data$searchperiod==0 & data$educ_primaryorless==1])

# Does not tell us anything, so ignore
prob10search_unknown = mean(data$benefits_week10[data$searchperiod==1 & data$educ_unknown==1])
prob10nosearch_unknown = mean(data$benefits_week10[data$searchperiod==0 & data$educ_unknown==1])
```

As Vocational, Prepvocational and Primaryorless go against the MIV assumption, we reduce the MIV to whether a person has a bachelor or not, and check again whether the MIV assumption holds:
```{r}
prob10search_NObachelormaster = mean(data$benefits_week10[data$searchperiod==1 & data$educ_bachelormaster==0])
prob10nosearch_NObachelormaster = mean(data$benefits_week10[data$searchperiod==0 & data$educ_bachelormaster==0])
```

Now, the MIV assumption holds as for both cases $d = \{0,1\}$, the probability of persons with a bachelor or master degree being still unemployed after 10 weeks is lower than the probability of persons without a bachelor or master degree. 
```{r}
prob_BachelorMaster = mean(data$educ_bachelormaster)
```

```{r}
# MIV for Benefits after 10 weeks

prob10nosearch_LB = (1-prob_BachelorMaster) * mean(data$benefits_week10[data$searchperiod==0 & data$educ_bachelormaster==0]) + prob_BachelorMaster * max(mean(data$benefits_week10[data$searchperiod==0 & data$educ_bachelormaster==0]), mean(data$benefits_week10[data$searchperiod==0 & data$educ_bachelormaster==1])) 
prob10nosearch_UB = (1-prob_BachelorMaster) * min(mean(data$benefits_week10[data$searchperiod==0 & data$educ_bachelormaster==0]), mean(data$benefits_week10[data$searchperiod==0 & data$educ_bachelormaster==1])) + prob_BachelorMaster * mean(data$benefits_week10[data$searchperiod==0 & data$educ_bachelormaster==1])

prob10search_LB = (1-prob_BachelorMaster) * mean(data$benefits_week10[data$searchperiod==1 & data$educ_bachelormaster==0]) + prob_BachelorMaster * max(mean(data$benefits_week10[data$searchperiod==1 & data$educ_bachelormaster==0]), mean(data$benefits_week10[data$searchperiod==1 & data$educ_bachelormaster==1])) 
prob10search_UB = (1-prob_BachelorMaster) * min(mean(data$benefits_week10[data$searchperiod==1 & data$educ_bachelormaster==0]), mean(data$benefits_week10[data$searchperiod==1 & data$educ_bachelormaster==1])) + prob_BachelorMaster * mean(data$benefits_week10[data$searchperiod==1 & data$educ_bachelormaster==1])

MIV10_LB = prob10search_LB - prob10nosearch_LB

MIV10_UB = prob10search_UB - prob10nosearch_UB
```

```{r}
# MIV for Benefits after 30 weeks

prob30nosearch_LB = (1-prob_BachelorMaster) * mean(data$benefits_week30[data$searchperiod==0 & data$educ_bachelormaster==0]) + prob_BachelorMaster * max(mean(data$benefits_week30[data$searchperiod==0 & data$educ_bachelormaster==0]), mean(data$benefits_week30[data$searchperiod==0 & data$educ_bachelormaster==1])) 
prob30nosearch_UB = (1-prob_BachelorMaster) * min(mean(data$benefits_week30[data$searchperiod==0 & data$educ_bachelormaster==0]), mean(data$benefits_week30[data$searchperiod==0 & data$educ_bachelormaster==1])) + prob_BachelorMaster * mean(data$benefits_week30[data$searchperiod==0 & data$educ_bachelormaster==1])

prob30search_LB = (1-prob_BachelorMaster) * mean(data$benefits_week30[data$searchperiod==1 & data$educ_bachelormaster==0]) + prob_BachelorMaster * max(mean(data$benefits_week30[data$searchperiod==1 & data$educ_bachelormaster==0]), mean(data$benefits_week30[data$searchperiod==1 & data$educ_bachelormaster==1])) 
prob30search_UB = (1-prob_BachelorMaster) * min(mean(data$benefits_week30[data$searchperiod==1 & data$educ_bachelormaster==0]), mean(data$benefits_week30[data$searchperiod==1 & data$educ_bachelormaster==1])) + prob_BachelorMaster * mean(data$benefits_week30[data$searchperiod==1 & data$educ_bachelormaster==1])

MIV30_LB = prob30search_LB - prob30nosearch_LB

MIV30_UB = prob30search_UB - prob30nosearch_UB
```

```{r}
# Dataframe with all bounds for benefits after 10 weeks
Model <- c('No Assumptions', 'Roy Model', 'MTS', 'MTR', 'MTS+MTR', 'MIV')
Lowerbound <- c(Noass10_LB, Roy10_LB, MTS10_LB, MTR10_LB, MTSMTR10_LB, MIV10_LB)
Upperbound <- c(Noass10_UB, Roy10_UB, MTS10_UB, MTR10_UB, MTSMTR10_UB, MIV10_UB)

bounds10 <- data.frame(Model, Lowerbound, Upperbound)
bounds10

# Dataframe with all bounds for benefits after 30 weeks
Model <- c('No Assumptions', 'Roy Model', 'MTS', 'MTR', 'MTS+MTR', 'MIV')
Lowerbound <- c(Noass30_LB, Roy30_LB, MTS30_LB, MTR30_LB, MTSMTR30_LB, MIV30_LB)
Upperbound <- c(Noass30_UB, Roy30_UB, MTS30_UB, MTR30_UB, MTSMTR30_UB, MIV30_UB)

bounds30 <- data.frame(Model, Lowerbound, Upperbound)
bounds30
```
```{r}
bounds10$Model  = factor(bounds10$Model, levels=c("No Assumptions", "Roy Model", "MTS", "MTR", "MTS+MTR", "MIV"))
library(ggplot2)
library(reshape2)
d <- melt(bounds10, id.vars="Model", value.name = "TreatmentEffect")

# Everything on the same plot
ggplot(d, aes(Model, TreatmentEffect, col=variable)) + 
  geom_line( color="grey") +
  geom_point(shape="-", color="#69b3a2", size=15) +
  stat_smooth() +
  geom_hline(yintercept=0)
```


```{r 7, results='asis'}
m10b <- lm(benefits_week10 ~ searchperiod + educ_bachelormaster + educ_prepvocational + educ_unknown + educ_vocational, data = data)
m30a <- lm(benefits_week30 ~ searchperiod + educ_bachelormaster + educ_prepvocational + educ_unknown + educ_vocational, data = data)

stargazer(m10a, m10b, m30a, m30b,type = "latex",
         title = "LPM for Benefits", header = FALSE)
```

